Lessons learned: creating tutorials teaching Agent-Based Modelling for Archaeologists

### 2024-09-18 paper.Rmd
Figure and language corrections

The surveys were analysed using R [@rcoreteam2023]. The following packages were used for analyses: ggplot2 [@wickham2016], dplyr [@wickham2023a], tidyr [@wickham2023], forcats[@wickham2023b], lubridate [@grolemund2011] and stringr [@wickham2023c]. The date and the code is available at <https://github.com/ABMArchaeologists/ABMA_paper> and (... insert Zenodo reference...)

@@

```{r nationality, fig.cap="The nationality of the participants that filled in the survey.", fig.height=5, echo=FALSE}
ggplot(pre_surveys, aes(x = `What.is.you.nationality....Nationality...Text`)) +
  geom_bar() +  coord_flip() + xlab("Nationality") + scale_x_discrete(limits=rev)

@@

```{r, abm-knowledge, fig.cap="Respondents knowledge and experience with ABM.", fig.show="hold", out.width="50%", echo=FALSE}
pre_surveys %>% 
  filter(`Do.you.know.what.Agent.Based.Modeling.is.`!="Not shared") %>% 
  ggplot(aes(x = `Do.you.know.what.Agent.Based.Modeling.is.`)) +
    geom_bar(na.rm = TRUE) + xlab("I know what ABM is") + facet_grid(~Workshop, labeller = event_names)
pre_surveys %>%
  select(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`) %>%
  filter(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`!="") %>%
  separate_rows(Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice, sep = ',\\s*') %>%
  group_by(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`) %>%
  ggplot(aes(x=`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`)) + geom_bar() +
  xlab("ABM Software known") + coord_flip()
```

@@

We also asked the participants about their intentions in terms of ABM and their professional life. The majority of the respondents - 88%, wanted to apply ABM in the future in their research (Figure \@ref(fig:future-abm)). Some of the respondents were already using ABM in their work. A large group (around 20-30%) was not sure yet how to apply ABM and wanted to read more on the subject or play a bit with the possibilities. Many respondents also shared the context to which they were thinking to apply ABM mentioning topics such as movement of people or goods over land or water, sometimes in relation to trade or other distribution mechanisms. Others thought of demography, social networks, migration or settlement distributions patterns. The natural environment and the interaction with humans in the past was also mentioned frequently, often in relation to GIS and the question of how complimentary these two computational techniques are. The archaeological periods that the participants were interested in were very diverse, ranging from the Paleolithic to the Medieval period.

```{r, future-abm, fig.cap="The respondents reaction to the question if they thing that they will apply ABM in the future.", out.height="50%", echo=FALSE}
post_surveys %>%
  filter(!is.na(Do.you.think.that.you.will.apply.ABM.to.your.own.research..If.yes..can.you.elaborate....Selected.Choice)) %>%
  ggplot(aes(x = 1, fill = `Do.you.think.that.you.will.apply.ABM.to.your.own.research..If.yes..can.you.elaborate....Selected.Choice`)) +
  geom_bar(na.rm = TRUE) + coord_polar(theta = "y", start=0) + theme_void() +
  scale_fill_discrete(name = paste0("Will apply ABM in the future (n=", sum(!is.na(post_surveys$Do.you.think.that.you.will.apply.ABM.to.your.own.research..If.yes..can.you.elaborate....Selected.Choice)), ")"),
                      labels = c("No", "Yes"))
```

>>>>

The surveys were analysed using R [@rcoreteam2023]. The following packages were used for analyses: ggplot2 [@wickham2016], dplyr [@wickham2023a], tidyr [@wickham2023], forcats[@wickham2023b], lubridate [@grolemund2011] and stringr [@wickham2023c]. For the sake of reproducibility [@marwick2017] this paper was written in R Markdown (<https://rmarkdown.rstudio.com/>), with data and code available at <https://github.com/ABMArchaeologists/ABMA_paper> (... insert Zenodo reference...)

@@

```{r nationality, fig.cap="The nationality of the participants that filled in the survey.", fig.height=5, echo=FALSE}
pre_surveys %>% 
  group_by(`What.is.you.nationality....Nationality...Text`) %>% 
  mutate(count_nationality = n()) %>%
  ggplot(aes(x=reorder(`What.is.you.nationality....Nationality...Text`, count_nationality))) + geom_bar() +  coord_flip() + xlab("Nationality")

@@

```{r, abm-knowledge, fig.cap="Respondents knowledge of ABM per event knowledge of the various softwares for ABM, before participating in the workshops.", fig.show="hold", out.width="50%", echo=FALSE}
pre_surveys %>%
  filter(`Do.you.know.what.Agent.Based.Modeling.is.`!="Not shared") %>%
  select(Do.you.know.what.Agent.Based.Modeling.is., Workshop) %>%
  group_by(`Do.you.know.what.Agent.Based.Modeling.is.`, Workshop) %>%
  mutate(count_response = n()) %>%
  distinct() %>%
  ggplot(aes(fill = `Do.you.know.what.Agent.Based.Modeling.is.`, x=Workshop, y = count_response)) +
  geom_bar(na.rm = TRUE, position="fill", stat="identity") +
  geom_text(aes(label=count_response), position = position_fill(vjust = 0.5), size = 2.5, colour = "grey41") + scale_x_discrete(labels = c("CAA", "EAA", "CAA-DE-NLFl", "Reuvensdagen", "CAA_UK", "Leiden", "Aarhus")) +
  xlab("") + scale_fill_discrete(name = "ABM known") + ylab("Percentage") +  scale_y_continuous(labels = scales::label_percent(scale = 100, prefix = "", suffix = "")) + coord_flip()
pre_surveys %>%
  select(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`) %>%
  filter(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`!="") %>%
  separate_rows(Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice, sep = ',\\s*') %>%
  group_by(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`) %>%
    mutate(count_software = n()) %>%
  ggplot(aes(x=reorder(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`, count_software))) + geom_bar() + scale_x_discrete(labels = ) + xlab("ABM Software known") + coord_flip()
```

@@

We also asked the participants about their intentions in terms of ABM and their professional life. The majority of the respondents - 88%, wanted to apply ABM in the future in their research. Some of the respondents were already using ABM in their work. A large group (around 20-30%) was not sure yet how to apply ABM and wanted to read more on the subject or play a bit with the possibilities. Many respondents also shared the context to which they were thinking to apply ABM mentioning topics such as movement of people or goods over land or water, sometimes in relation to trade or other distribution mechanisms. Others thought of demography, social networks, migration or settlement distributions patterns. The natural environment and the interaction with humans in the past was also mentioned frequently, often in relation to GIS and the question of how complimentary these two computational techniques are. The archaeological periods that the participants were interested in were very diverse, ranging from the Paleolithic to the Medieval period.

### 2024-09-18 paper.tex
Figure and language corrections

The surveys were analysed using R (R Core Team 2023). The following packages were used for analyses: ggplot2 (Wickham 2016), dplyr (Wickham et al. 2023), tidyr (Wickham, Vaughan and Girlich 2023), forcats(Wickham 2023a), lubridate (Grolemund and Wickham 2011) and stringr (Wickham 2023b). The date and the code is available at \url{https://github.com/ABMArchaeologists/ABMA_paper} and (\ldots{} insert Zenodo reference\ldots)

@@

\begin{figure}
\includegraphics[width=0.5\linewidth]{paper_files/figure-latex/abm-knowledge-1} \includegraphics[width=0.5\linewidth]{paper_files/figure-latex/abm-knowledge-2} \caption{Respondents knowledge and experience with ABM.}\label{fig:abm-knowledge}
\end{figure}

@@

We also asked the participants about their intentions in terms of ABM and their professional life. The majority of the respondents - 88\%, wanted to apply ABM in the future in their research (Figure \ref{fig:future-abm}). Some of the respondents were already using ABM in their work. A large group (around 20-30\%) was not sure yet how to apply ABM and wanted to read more on the subject or play a bit with the possibilities. Many respondents also shared the context to which they were thinking to apply ABM mentioning topics such as movement of people or goods over land or water, sometimes in relation to trade or other distribution mechanisms. Others thought of demography, social networks, migration or settlement distributions patterns. The natural environment and the interaction with humans in the past was also mentioned frequently, often in relation to GIS and the question of how complimentary these two computational techniques are. The archaeological periods that the participants were interested in were very diverse, ranging from the Paleolithic to the Medieval period.

\begin{figure}
\includegraphics[height=0.5\textheight]{paper_files/figure-latex/future-abm-1} \caption{The respondents reaction to the question if they thing that they will apply ABM in the future.}\label{fig:future-abm}
\end{figure}

@@

>>>>

The surveys were analysed using R (R Core Team 2023). The following packages were used for analyses: ggplot2 (Wickham 2016), dplyr (Wickham et al. 2023), tidyr (Wickham, Vaughan and Girlich 2023), forcats(Wickham 2023a), lubridate (Grolemund and Wickham 2011) and stringr (Wickham 2023b). For the sake of reproducibility (Marwick 2017) this paper was written in R Markdown (\url{https://rmarkdown.rstudio.com/}), with data and code available at \url{https://github.com/ABMArchaeologists/ABMA_paper} (\ldots{} insert Zenodo reference\ldots)

@@

\begin{figure}
\includegraphics[width=0.5\linewidth]{paper_files/figure-latex/abm-knowledge-1} \includegraphics[width=0.5\linewidth]{paper_files/figure-latex/abm-knowledge-2} \caption{Respondents knowledge of ABM per event knowledge of the various softwares for ABM, before participating in the workshops.}\label{fig:abm-knowledge}
\end{figure}

@@

We also asked the participants about their intentions in terms of ABM and their professional life. The majority of the respondents - 88\%, wanted to apply ABM in the future in their research. Some of the respondents were already using ABM in their work. A large group (around 20-30\%) was not sure yet how to apply ABM and wanted to read more on the subject or play a bit with the possibilities. Many respondents also shared the context to which they were thinking to apply ABM mentioning topics such as movement of people or goods over land or water, sometimes in relation to trade or other distribution mechanisms. Others thought of demography, social networks, migration or settlement distributions patterns. The natural environment and the interaction with humans in the past was also mentioned frequently, often in relation to GIS and the question of how complimentary these two computational techniques are. The archaeological periods that the participants were interested in were very diverse, ranging from the Paleolithic to the Medieval period.

@@

\leavevmode\vadjust pre{\hypertarget{ref-marwick2017}{}}%
Marwick, B. 2017 Computational Reproducibility in Archaeological Research: Basic Principles and a Case Study of Their Implementation. \emph{Journal of Archaeological Method and Theory} 24(2): 424--450. DOI: https://doi.org/\href{https://doi.org/10.1007/s10816-015-9272-9}{10.1007/s10816-015-9272-9}.
